---
title: "CMAP"
author: "Luke"
date: "January 18, 2016"
output: html_document
---
#Load Packages
##std stats libraries
```{r,results="hide", warning=FALSE, message=FALSE}
library(psych)
library(stats)
library(plyr)
library(MASS)
install.packages('leaps')
library(leaps)
library(car)
install.packages('corrplot')
library(corrplot)
library(ggplot2)
library(cluster)   #for clusplot
library(data.table)
library(knitr)
library(dplyr)
```
##Libraries required for Geo Map Tracts plot
```{r,results="hide", warning=FALSE, message=FALSE}
library(rgdal)
library(scales)
library(ggmap)
install.packages('Cairo')
library(Cairo)
install.packages("gpclib", type="source") # RUN THIS IF YOU GET AN ERROR ON 'fortify' BELOW
library(rgeos)
library(maptools)
library(data.table)
```
#Preprocessing
##load data
###The CSV_file is the meta data(MasterFile)
```{r}
setwd("~/Desktop/housing projects/R/cmapdata")
CSV_file <- read.csv("MasterFile_ExcludeTracts_1980.csv", head=TRUE, sep=",", stringsAsFactors = F)
```
##Split data
###Splitting the data for future analyze 
###1.cmap2013;2.data change. All in format of dataframe. Then Skip the total population and total HouseUnits
```{r}
cmap2013.df = CSV_file[, c(1, 6, 8, 10, 12:22, 24, 26, 28, 30:32, 34, 36, 38, 40,
              42, 44, 46:47, 53:57, 59, 61, 63:64, 67, 69, 71, 73:76, 84, 86)]  

chg00_13.df = CSV_file[, c(1, 3, 5, 7, 9, 11, 23, 25, 27, 29, 30:31, 33, 35, 37, 39,
              41, 43, 45, 48:52, 58, 60, 62, 65:66, 68, 70, 72, 77:83, 85, 87)]
##Skip Y2013 Total Population and Total HouseUnits in Columns 2 & 4
cmap2013.df = CSV_file[, c(6, 8, 10, 12:22, 24, 26, 28, 30:32, 34, 36, 38, 40,
              42, 44, 46:47, 53:57, 59, 61, 63:64, 67, 69, 71, 73:76,84, 86)]  

chg00_13.df = CSV_file[, c(3, 5, 7, 9, 11, 23, 25, 27, 29, 30:31, 33, 35, 37, 39,
              41, 43, 45, 48:52, 58, 60, 62, 65:66, 68, 70, 72, 77:83, 85, 87:90)] 

#???????remove desire variables from  datasets ?What means desire?
cmap2013.df$Y13_MedGrRent_Scaled <- NULL
cmap2013.df$Y13_PCT_HUnits_CostBurden_LT30 <- NULL
cmap2013.df$Y14PCT_Renter_Moved1980to1999 <- NULL
chg00_13.df$Chg_PCT_MedGrossRent <- NULL
chg00_13.df$OwnerChange <- NULL
chg00_13.df$HUD_Housing_Pct_of_HHIncome_hh_h <- NULL
chg00_13.df$HUD_Transportation_Pct_of_HHIncome_hh_t <- NULL
```

Cluster-PAM
#Using PAM cluster to make segmentation of tracts based on variables we have
```{r}
km13.data = data.table(cmap2013.df)
kmchg.data = data.table(chg00_13.df)

#??????????????????????set all NA values to 0? or -1
km13.data[is.na(km13.data)] = -1
kmchg.data[is.na(kmchg.data)] = -1

km13.scaled = scale(km13.data)
kmchg.scaled = scale(kmchg.data)
#-- try clusplot from pkg CLUSTER
par(mar=c(5, 4, 4, 2))   #default

##Selecting kmchg to analyze
fitpam1 <- pam(kmchg.scaled, 8, metric = "euclidean")
fitpam1$clusinfo

#Intergrated Tract num with pam cluster and pam data
##1.Transfer pam.clusters to data frame
##2.Assign CensusTract column from CSV_file
##3.Assign pam.data <- kmchg.data to data frame
##4.Intergrated two df(clusters idx, kmchgdata)
##We have a data frame called pam.clusters.(The chg-data with pam cluster idx)
pam.clusters <- as.data.frame(fitpam1$clustering)
pam.clusters$CensusTract = CSV_file$CensusTract
pam.data = as.data.frame(kmchg.data)
pam.data$CensusTract = CSV_file$CensusTract
pam.clusters <- merge(pam.clusters,pam.data,by='CensusTract')
names(pam.clusters)[2] <- "ClusterNum"
```

Boxplotting for each Variables across different clusters
```{r}
VariMaxData = data.table(pam.clusters[,2:41])
#Create df for each cluster exclude column ClusterNum
Varidata1 = as.data.frame( VariMaxData[ClusterNum == 1, -c("ClusterNum"), with=FALSE ] )
Varidata2 = as.data.frame( VariMaxData[ClusterNum == 2, -c("ClusterNum"), with=FALSE ] )
Varidata3 = as.data.frame( VariMaxData[ClusterNum == 3, -c("ClusterNum"), with=FALSE ] )
Varidata4 = as.data.frame( VariMaxData[ClusterNum == 4, -c("ClusterNum"), with=FALSE ] )
Varidata5 = as.data.frame( VariMaxData[ClusterNum == 5, -c("ClusterNum"), with=FALSE ] )
Varidata6 = as.data.frame( VariMaxData[ClusterNum == 6, -c("ClusterNum"), with=FALSE ] )
Varidata7 = as.data.frame( VariMaxData[ClusterNum == 7, -c("ClusterNum"), with=FALSE ] )
Varidata8 = as.data.frame( VariMaxData[ClusterNum == 8, -c("ClusterNum"), with=FALSE ] )

#set the correct column to start plotting X variables
i = 1
col_num <- length(Varidata1)

# loop thru columns and plot boxplot of each X variable
while (i <= col_num){
boxplot(Varidata1[,i], Varidata2[,i], Varidata3[,i], Varidata4[,i], Varidata5[,i], Varidata6[,i],
        Varidata7[,i], Varidata8[,i],  
        horizontal=TRUE,
        main = names(Varidata1)[i],
        names=c("Clus1", "Clus2", "Clus3", "Clus4", "Clus5", "Clus6", "Clus7",
                "Clus8"     )   )

CurrVariable = names(Varidata1)[i]
print(CurrVariable)
print('Cluster1 Summary')
print(summary(Varidata1[,i]))
print('Cluster2 Summary')
print(summary(Varidata2[,i]))
print('Cluster3 Summary')
print(summary(Varidata3[,i]))
print('Cluster4 Summary')
print(summary(Varidata4[,i]))
print('Cluster5 Summary')
print(summary(Varidata5[,i]))
print('Cluster6 Summary')
print(summary(Varidata6[,i]))
print('Cluster7 Summary')
print(summary(Varidata7[,i]))
print('Cluster8 Summary')
print(summary(Varidata8[,i]))

i = i + 1
}
```

#Plotting County, Tracts fill color by Clusters
```{r}
# tract shapefile
##Q: What is a tract shapefile?
##A: Main job is to present id, long, lat.
tract.shapefile <- read.csv("tract_shapefile_2010.csv", head=TRUE, sep=",", stringsAsFactors = F)
#-- column of Tract IDs must be of type CHAR for join, then rename the CensusTract: With 1400000US
pam.clusters$CensusTract <- as.character(pam.clusters$CensusTract) 
pam.clusters$CensusTract = paste("1400000US", pam.clusters$CensusTract, sep = "")

#---- Chicago Tract Map PLOT for chg Now Clusters  ------
##First layer -------Tracts
plotData1<- left_join(tract.shapefile, pam.clusters, by = c("id" = "CensusTract")) #join data to the map data
yourMap <- ggplot() +
    geom_polygon(data = plotData1, aes(x=long, y=lat, group = group, fill = factor(ClusterNum)), colour = "black", size = 0.15)
print(yourMap)


# county shapefile
##Q: What is a county shapefile?
##A: Main job is to present id, long, lat of county Boundaries
county.shapefile <- read.csv("county_shapefile_2010.csv", head=TRUE, sep=",", stringsAsFactors = F)
# countyInfo
tract_county <- read.csv("CensusTract2013_to_CountyIDs_forShapeFiles.csv", head=TRUE, sep=",", stringsAsFactors = F)

tract_county$CensusTract2013 = as.character(tract_county$CensusTract2013)
tract_county$CensusTract = paste("1400000US", tract_county$CensusTract2013, sep = "")
tract_county$countyidChar <- as.character(tract_county$CountyID)

#Integrating three datasets pam.cluster <---> tract_county <---> county.shapefile
Middledata <- left_join(pam.clusters, tract_county, by = c("CensusTract" = "CensusTract")) 
Middledata$countyidChar = as.character(Middledata$countyidChar)
county.shapefile$id = as.character(county.shapefile$id)
plotData2 <- left_join(county.shapefile, Middledata, by = c("id" = "countyidChar"))

#Second layer ---- tract + county
yourMap2 <- yourMap+
    geom_polygon(data = plotData2, aes(x = long, y = lat, group = group), colour = "white", fill = NA)
print(yourMap2)
```

#Punctuation Layer
```{r}
# load Gazetteer file
Gazetteer <- read.table("Illinois Gazetteer Files 2010.txt", head=TRUE, stringsAsFactors = F)
Gazetteer.geo = Gazetteer[,c("GEOID","INTPTLAT","INTPTLONG")]
colnames(Gazetteer.geo) <- c("id", "lat", "long")

#Selecting Variables interested in 
#As for Chg_Educate_HighSchool_orLess
# bin Chg_Educate_HighSchool_orLess into 4 groups
#NA values exist here as there is a miserable tract

hist(pam.clusters$Chg_Educate_HighSchool_orLess)
quantileL = quantile(pam.clusters$Chg_Educate_HighSchool_orLess, names = FALSE,na.rm=TRUE)

# create new variable "Cha_Chg_Educate_HighSchool_orLess"
pam.clusters$Cha_Chg_Educate_HighSchool_orLess = 1
pam.clusters$Cha_Chg_Educate_HighSchool_orLess[which(CSV_file$Chg_Educate_HighSchool_orLess>quantileL[2])] = 2
pam.clusters$Cha_Chg_Educate_HighSchool_orLess[which(CSV_file$Chg_Educate_HighSchool_orLess>quantileL[3])] = 3
pam.clusters$Cha_Chg_Educate_HighSchool_orLess[which(CSV_file$Chg_Educate_HighSchool_orLess>quantileL[4])] = 4
hist(pam.clusters$Cha_Chg_Educate_HighSchool_orLess)
# convert Cha_Chg_Educate_HighSchool_orLess into factor
pam.clusters$Cha_Chg_Educate_HighSchool_orLess = as.factor(pam.clusters$Cha_Chg_Educate_HighSchool_orLess)

Gazetteer.geo$id = paste("1400000US", Gazetteer.geo$id, sep = "")
plotDataPunc <- left_join(pam.clusters, Gazetteer.geo, by = c("CensusTract" = "id")) #join data to the map data
#----tract + county + punctuation
yourMap3 <- yourMap2+
    geom_point(plotDataPunc, aes(x=long, y=lat,shape=factor(Cha_Chg_Educate_HighSchool_orLess)))
print(yourMap3)
```