---
title: "mappingmethod"
author: "Luke"
date: "January 19, 2016"
output: html_document
---
#LOAD PACKAGES
```{r,results="hide"}
library(psych)
library(stats)
library(plyr)
library(MASS)
library(leaps)
library(car)
library(corrplot)
library(ggplot2)
library(cluster)   #for clusplot
library(data.table)
library(knitr)
library(dplyr)
```
##Libraries required for Geo Map Tracts plot
```{r,results="hide"}
library(rgdal)
library(scales)
library(ggmap)
#install.packages('Cairo')
#library(Cairo)
install.packages("gpclib", type="source") # RUN THIS IF YOU GET AN ERROR ON 'fortify' BELOW
library(rgeos)
library(maptools)
library(data.table)
```
#PRELOAD AND PROCESSING
```{r}
###The CSV_file is the meta data(MasterFile)
setwd("~/Desktop/housing projects/R/cmapdata")
CSV_file <- read.csv("MasterFile_ExcludeTracts_1980.csv", head=TRUE, sep=",", stringsAsFactors = F)
###Splitting the data for future analyze 
#1.cmap2013
#2.data change
####All in format of dataframe. Then Skip the total population and total HouseUnits

cmap2013.df = CSV_file[, c(1, 6, 8, 10, 12:22, 24, 26, 28, 30:32, 34, 36, 38, 40,
              42, 44, 46:47, 53:57, 59, 61, 63:64, 67, 69, 71, 73:76, 84, 86)]  

chg00_13.df = CSV_file[, c(1, 3, 5, 7, 9, 11, 23, 25, 27, 29, 30:31, 33, 35, 37, 39,
              41, 43, 45, 48:52, 58, 60, 62, 65:66, 68, 70, 72, 77:83, 85, 87)]
##Skip Y2013 Total Population and Total HouseUnits in Columns 2 & 4
cmap2013.df = CSV_file[, c(6, 8, 10, 12:22, 24, 26, 28, 30:32, 34, 36, 38, 40,
              42, 44, 46:47, 53:57, 59, 61, 63:64, 67, 69, 71, 73:76,84, 86)]  

chg00_13.df = CSV_file[, c(3, 5, 7, 9, 11, 23, 25, 27, 29, 30:31, 33, 35, 37, 39,
              41, 43, 45, 48:52, 58, 60, 62, 65:66, 68, 70, 72, 77:83, 85, 87:90)] 

#???????remove desire variables from  datasets ?What means desire?
cmap2013.df$Y13_MedGrRent_Scaled <- NULL
cmap2013.df$Y13_PCT_HUnits_CostBurden_LT30 <- NULL
cmap2013.df$Y14PCT_Renter_Moved1980to1999 <- NULL
chg00_13.df$Chg_PCT_MedGrossRent <- NULL
chg00_13.df$OwnerChange <- NULL
chg00_13.df$HUD_Housing_Pct_of_HHIncome_hh_h <- NULL
chg00_13.df$HUD_Transportation_Pct_of_HHIncome_hh_t <- NULL

#Clustering
##Cluster-PAM
##Using PAM cluster to make segmentation of tracts based on variables we have
###km13.data & kmchg.data
#####Analyzing 2013 data or the chg data?
km13.data = data.table(cmap2013.df)
kmchg.data = data.table(chg00_13.df)
#??????????????????????set all NA values to 0? or -1
km13.data[is.na(km13.data)] = -1
kmchg.data[is.na(kmchg.data)] = -1
km13.scaled = scale(km13.data)
kmchg.scaled = scale(kmchg.data)
#-- try clusplot from pkg CLUSTER
par(mar=c(5, 4, 4, 2))   #default
```


#Over view
##1.Create cluster data(cluster data------data need to be analyze with cluster idx)
##2.Bin Non-numeric Variables 
##3.Concatenating shape files(2) punctiation file(1) to cluster data 
##4.Generate tract based, county based, punctuation based mapping---layer by layer

##1.Generate cluster data
###The input cluster data is kmchg or km13 both scaled and k needs to be specified
```{r}
gencludata <- function(data,k){
    fitpam <- pam(data, k, metric = "euclidean")
    clusteridx <- as.data.frame(fitpam$clustering)
    clusteridx$CensusTract <- CSV_file$CensusTract
    cmbindata <- as.data.frame(data)
    cmbindata$CensusTract <- CSV_file$CensusTract
    integratedata <- merge(clusteridx,cmbindata,by='CensusTract')
    names(integratedata)[2]<-"ClusterNum"
    return(integratedata)
}
```

##2.Bin Non-numeric Variables 
###More Variables contains in bin rmd file
```{r}
integratedata <- gencludata(kmchg.scaled,8)

hist(integratedata$Chg_Educate_HighSchool_orLess)
quantileL = quantile(integratedata$Chg_Educate_HighSchool_orLess, names = FALSE)

# always put an order number in front of letter! This will help us when we convert it
integratedata$Cha_Chg_Educate_HighSchool_orLess = "1-Low"
integratedata$Cha_Chg_Educate_HighSchool_orLess[which(integratedata$Chg_Educate_HighSchool_orLess>quantileL[2])] = "2-Median"
integratedata$Cha_Chg_Educate_HighSchool_orLess[which(integratedata$Chg_Educate_HighSchool_orLess>quantileL[3])] = "3-High"
integratedata$Cha_Chg_Educate_HighSchool_orLess[which(integratedata$Chg_Educate_HighSchool_orLess>quantileL[4])] = "4-Very High"
# check result
hist(as.numeric(substr(integratedata$Cha_Chg_Educate_HighSchool_orLess,0,1)))

# convert into factor
integratedata$Cha_Chg_Educate_HighSchool_orLess = as.factor(integratedata$Cha_Chg_Educate_HighSchool_orLess)
```

##3.Concatenating shape files(2) punctiation file(1) to cluster data 
```{r}
#integratedata <- gencludata(kmchg.scaled,8)
#Load 3 shape files
tract.shapefile <- read.csv("tract_shapefile_2010.csv", head=TRUE, sep=",", stringsAsFactors = F)
county.shapefile <- read.csv("county_shapefile_2010.csv", head=TRUE, sep=",", stringsAsFactors = F)
tract_county <- read.csv("CensusTract2013_to_CountyIDs_forShapeFiles.csv", head=TRUE, sep=",", stringsAsFactors = F)
Gazetteer <- read.table("Illinois Gazetteer Files 2010.txt", head=TRUE, stringsAsFactors = F)
Gazetteer.geo = Gazetteer[,c("GEOID","INTPTLAT","INTPTLONG")]
colnames(Gazetteer.geo) <- c("id", "lat", "long")
#Concatenating
#1 Tract
integratedata$CensusTract <- as.character(integratedata$CensusTract) 
integratedata$CensusTract = paste("1400000US", integratedata$CensusTract, sep = "")
plotData1<- left_join(tract.shapefile, integratedata, by = c("id" = "CensusTract")) 

#2 Tract+County
#Integrating three datasets integratedata <---> tract_county <---> county.shapefile
tract_county$CensusTract2013 = as.character(tract_county$CensusTract2013)
tract_county$CensusTract = paste("1400000US", tract_county$CensusTract2013, sep = "")
tract_county$countyidChar <- as.character(tract_county$CountyID)

Middledata <- left_join(integratedata, tract_county, by = c("CensusTract" = "CensusTract")) 
Middledata$countyidChar = as.character(Middledata$countyidChar)
county.shapefile$id = as.character(county.shapefile$id)
plotData2 <- left_join(county.shapefile, Middledata, by = c("id" = "countyidChar"))

#3 Punctuation
Gazetteer.geo$id = paste("1400000US", Gazetteer.geo$id, sep = "")
plotDataPunc <- left_join(integratedata, Gazetteer.geo, by = c("CensusTract" = "id"))
```

##4.Generate tract based, county based, punctuation based mapping---layer by layer
```{r}
yourMap <- ggplot() +
    geom_polygon(data = plotData1, aes(x=long, y=lat, group = group, fill = factor(ClusterNum)), colour = "black", size = 0.2)
#print(yourMap)

yourMap2 <- yourMap+
    geom_polygon(data = plotData2, aes(x = long, y = lat, group = group), colour = "white", fill = NA)
#print(yourMap2)
```

```{r}
yourMap3 <- yourMap2+
    geom_point(data = plotDataPunc, aes(x=long, y=lat,shape=factor(Cha_Chg_Educate_HighSchool_orLess)))
print(yourMap3)
```



