---
title: "luke analysis-CMAPproject"
author: "Luke"
date: "February 12, 2016"
output: html_document
---

Packages
```{r-setup files & libraries}
library(psych)
library(stats)
library(plyr)
library(MASS)
library(leaps)
library(car)
library(corrplot)
library(ggplot2)
library(cluster)   #for clusplot
library(data.table)
library(knitr)
library(dplyr)
library(VIM)
#install.packages('clValid')
library(clValid)
library(xlsx)
```

Load Data
```{r}
path = '~/Desktop/housing projects/R/data'
setwd(path)
CSV_file <- read.csv('MasterFile_1980.csv', head=TRUE, sep=",", stringsAsFactors = F)  
origHomeicm_file <- read.csv('HomeValueIncome_UNscaled.csv',sep=',',head=TRUE,stringsAsFactors = F)
orighomeval.df<-origHomeicm_file[,c(1:7)]
```

----- Split CSV_file 1 into separate DFs ---------
Skip Y2013 Total Population and Total HouseUnits in Columns 2 & 4
Ignore Census Tract for initial analysis like COrrelation
Skip Y2013 Total Population and Total HouseUnits in Columns 2 & 4
```{r-split files preprocessing}
cmap2013.df = CSV_file[, c(1, 6, 8, 10, 12:22, 24, 26, 28, 30:32, 34, 36, 38, 40,
              42, 44, 46:47, 53:57, 59, 61, 63:64, 67, 69, 71, 73:76,84, 86)]  
chg00_13.df = CSV_file[, c(1, 3, 5, 7, 9, 11, 23, 25, 27, 29, 30:31, 33, 35, 37, 39,
              41, 43, 45, 48:52, 58, 60, 62, 65:66, 68, 70, 72, 77:83, 85, 87:90)]    
```
--- Full 46 Vars with extra Y2000 variables added -----------
remove desire variables from  datasets 
```{r-split files preprocessing}
#cmap2013
cmap2013.df$Y13_MedGrRent_Scaled <- NULL
cmap2013.df$Y13_PCT_HUnits_CostBurden_LT30 <- NULL
cmap2013.df$Y14PCT_Renter_Moved1980to1999 <- NULL
cmap2013.df$Y14_PCT_OtherRace_Alone <- NULL
cmap2013.df$Y14_PCT_Not_Hispanic_or_Latino_Total <- NULL
cmap2013.df$Y14_PCT_Two_or_More_Races <- NULL
cmap2013.df$Y13_PCT_OwnerCount <- NULL
cmap2013.df$Y13_MedGrRent_Scaled <- NULL
cmap2013.df$Scaled_PopDensity2013 <- NULL
cmap2013.df$Y13_PCT_HUnits_CostBurden_LT30 <- NULL
cmap2013.df$Y14PCT_Renter_Moved2000to2010_orLater  <- NULL
#Deal with MissingValues
cmap2013.missv = aggr(cmap2013.df, prop=FALSE, numbers=TRUE)
summary(cmap2013.missv)
#chg00_13
chg00_13.df$Chg_PCT_MedGrossRent <- NULL
chg00_13.df$OwnerChange <- NULL
chg00_13.df$Change_CostBurden_LTE30 <- NULL
chg00_13.df$Chg_Scaled_PopDensity <- NULL
chg00_13.df$HUD_Housing_Pct_of_HHIncome_hh_h <- NULL
chg00_13.df$HUD_Transportation_Pct_of_HHIncome_hh_t <- NULL
chg00_13.df$Chg_PCT_Not_Hispanic_or_Latino  <- NULL
chg00_13.df$Chg_PCT_OtherRace_Alone   <- NULL
chg00_13.df$Chg_PCT_Two_or_More_Races  <- NULL

chg00_13.missv = aggr(chg00_13.df, prop=FALSE, numbers=TRUE)
summary(chg00_13.missv)
```

```{r clustering PAM preprocessing}
km13.data = data.table(cmap2013.df)
kmchg.data = data.table(chg00_13.df)

km13.data = na.omit(km13.data)
km13.bkp = km13.data
km13.data$CensusTract <- NULL

kmgchg.nasbset = na.omit(kmchg.data, invert=TRUE)
kmchg.data = na.omit(kmchg.data)
kmchg.bkp = kmchg.data #kmchg backup
kmchg.data$CensusTract <- NULL
```

Seleting Clustering Methods
```{r}
######################################
kmchg.data <- data.frame(kmchg.data)##
######################################
intern <- clValid(kmchg.data[1:nrow(kmchg.data),],5:12,clMethods=c('hierarchical','kmeans','pam'),validation='internal')
getRanksWeights(intern)$ranks

stab <- clValid(kmchg.data[1:nrow(kmchg.data),],5:12,clMethods=c('hierarchical','kmeans','pam'),validation='stability')#Time consuming and not converge in 10 iterations
getRanksWeights(stab)$ranks

result <- clValid(kmchg.data[1:nrow(kmchg.data),],3:15,clMethods=c('kmeans','pam'),validation=c('internal'))
res <- getRanksWeights(result)
print(res$ranks,quote=FALSE)

#Internal 2:15 different method
result1 <- clValid(kmchg.data[1:nrow(kmchg.data),],2:15,clMethods=c('hierarchical'),validation=c('internal'))
res1 <- getRanksWeights(result1)
print(res1$ranks,quote=FALSE)

result2 <- clValid(kmchg.data[1:nrow(kmchg.data),],2:15,clMethods=c('kmeans'),validation=c('internal'))
res2 <- getRanksWeights(result2)
print(res2$ranks,quote=FALSE)

result3 <- clValid(kmchg.data[1:nrow(kmchg.data),],2:15,clMethods=c('pam'),validation=c('internal'))
res3 <- getRanksWeights(result3)
print(res3$ranks,quote=FALSE)

#Hierarchical clustering
d <- dist(kmchg.data,method='euclidean')
fit <-hclust(d,method='ward.D')
group<- cutree(fit,k=7)
rect.hclust(fit,k=7,border='red')
```

Producting Cluster index number
#Both pamindex and kmindex aim to catch the input data and parameters 
#then return the dataset with cluster index attached with original data
```{r clust index function}
pamindex <- function (data,k,distmethod){
    fit <- pam(data,k,metric = distmethod)
    pamclster = as.data.frame(fit$clustering)
    pamclster$CensusTract =  kmchg.bkp$CensusTract     
    pam.data = as.data.frame(kmchg.data)
    pam.data$CensusTract =  kmchg.bkp$CensusTract              
    pamclster <- merge(pamclster, pam.data, by="CensusTract")
    names(pamclster)[2] <- "ClusterNum"
    VariMaxData = pamclster 
    return (VariMaxData)
}
kmindex<- function(data,k){
    fit <- kmeans(data,k)
    kmclster = as.data.frame(fit$cluster)
    kmclster$CensusTract =  kmchg.bkp$CensusTract     
    km.data = as.data.frame(kmchg.data)
    km.data$CensusTract =  kmchg.bkp$CensusTract              
    kmclster <- merge(kmclster, km.data, by="CensusTract")
    names(kmclster)[2] <- "ClusterNum"
    VariMaxData = kmclster
    return (VariMaxData)
}
```

#Both anovatest function will return the variable doesn't pass the anova test with relative clustering #methond and the boxplot of each variable.
```{r ANOVA test variance confirm on N clusters of method M}
boxp1 <- function(data,i){
    data <- data.frame(data)
    p <- ggplot(data, aes(x=ClusterNum, y=data[,i])) +
        geom_boxplot(aes(group=ClusterNum,y =data[,i],color = factor(ClusterNum))) +
        scale_x_continuous(breaks = seq(1,8,by=1)) +
        theme(axis.text.x = element_text(angle=45, vjust=0.5)) +
        ylab(colnames(data)[i]) +
        ggtitle(paste('Boxplot for',colnames(data)[i] ,'across each cluster'))
    print (p)
}
anovatestpam <- function (data,k,distmethod){
    datause <- pamindex(data,k,distmethod)
    for (i in 3:ncol(datause)){
        anovtest = aov(datause[,i]~datause[,2], data=datause)
        p_value <- summary(anovtest)[[1]][["Pr(>F)"]][1]
        if (p_value > 0.05){
            print(c(i,colnames(datause)[i],p_value))
            boxp1(datause,i)
        }
    }
}

anovatestkm <- function (data,k){
    datause <- kmindex(data,k)
    for (i in 3:ncol(datause)){
        anovtest = aov(datause[,i]~datause[,2], data=datause)
        p_value <- summary(anovtest)[[1]][["Pr(>F)"]][1]
        if (p_value > 0.05){
            print(c(i,colnames(datause)[i],p_value))
            boxp1(datause,i)
        }
    }
}
```

```{r luke ploting data comparison among clusters}
boxp <- function(data1){
    data1 <- data.frame(data1)
    col_num <- ncol(data1)
    for (i in 2:col_num){
        p <- ggplot(data1, aes(x=ClusterNum, y=data1[,i])) +
            geom_boxplot(aes(group=ClusterNum,y =data1[,i],color = factor(ClusterNum))) +
            scale_x_continuous(breaks = seq(1,8,by=1)) +
            theme(axis.text.x = element_text(angle=45, vjust=0.5)) +
            ylab(colnames(data1)[i]) +
            ggtitle(paste('Boxplot for',colnames(data1)[i] ,'across each cluster'))
        print (p)
    }
}

vilop <- function(data1){
    data1 <- data.frame(data1)
    col_num <- ncol(data1)
    for (i in 2:col_num){
        p <- ggplot(data1, aes(x=ClusterNum, y=data1[,i])) +
            geom_violin(aes(group=ClusterNum,y =data1[,i],color = factor(ClusterNum))) +
            scale_x_continuous(breaks = seq(1,8,by=1)) +
            theme(axis.text.x = element_text(angle=45, vjust=0.5)) +
            ylab(colnames(data1)[i]) +
            ggtitle(paste('Boxplot for',colnames(data1)[i] ,'across each cluster'))
        print (p)
    }
}

histcpr <- function(data1){
    data1 <- data.frame(data1)
    col_num <- ncol(data1)
    for (i in 2:col_num){
        p <- ggplot(data1,aes(data1[,i],fill=factor(ClusterNum))) +
            geom_histogram(color = 'grey30',binwidth = 0.05) +
            facet_grid(ClusterNum~.) +
            xlab(colnames(data1)[i]) +
            ylab(colnames(data1)[i]) +
            ggtitle(paste('histogram for',colnames(data1)[i] ,'across each cluster'))
        print (p)
    }
}
```

```{r load geo basefile}
#Load shape file
tract.shapefile <- read.csv("tract_shapefile_2010.csv", head=TRUE, sep=",", stringsAsFactors = F)
county.shapefile <- read.csv("county_shapefile_2010.csv", head=TRUE, sep=",", stringsAsFactors = F)
tract_county <- read.csv("CensusTract2013_to_CountyIDs_forShapeFiles.csv", head=TRUE, sep=",", stringsAsFactors = F)
Gazetteer <- read.table("Illinois Gazetteer Files 2010.txt", head=TRUE, stringsAsFactors = F)
Gazetteer.geo = Gazetteer[,c("GEOID","INTPTLAT","INTPTLONG")]
colnames(Gazetteer.geo) <- c("id", "lat", "long")
```

```{r combine file by index-censustract}
tract_county$CensusTract2013 = as.character(tract_county$CensusTract2013)
tract_county$CensusTract = paste("1400000US", tract_county$CensusTract2013, sep = "")
tract_county$countyidChar <- as.character(tract_county$CountyID)
Gazetteer.geo$id = paste("1400000US", Gazetteer.geo$id, sep = "")

#input data is pamindex or kmindex function
#Give data a 1400000US index
reviseid_140US <- function(data){
    data$CensusTract <- as.character(data$CensusTract) 
    data$CensusTract = paste("1400000US", data$CensusTract, sep = "")
    return(data)
}

#Combination of cluster indexed data and tracts
#Tract level cluster data
Tractdata <- function(data){
    data <- reviseid_140US(data)
    Tractlevel<- left_join(tract.shapefile, data, by = c("id" = "CensusTract"))
    return(Tractlevel)
}

#input tractlevel data and concatenate with county-tract level index
middledata <- function(data){   
    data <- reviseid_140US(data)
    Middledata <- left_join(data, tract_county, by = c("CensusTract" = "CensusTract")) 
    Middledata$countyidChar = as.character(Middledata$countyidChar)
    return (Middledata)
}


Countydata <- function(data){
    Mdata <- middledata(data)
    county.shapefile$id = as.character(county.shapefile$id)
    CountyTractlevel <- left_join(county.shapefile, Mdata,by = c("id"="countyidChar" ))
    return (CountyTractlevel)
    #write.csv(CountyTractlevel,path) return 20.52GB file......
}

buildpunc <- function(data){
    data <- reviseid_140US(data)
    Punc_data <- left_join(data,Gazetteer.geo,by=c('CensusTract' = 'id'))
    return (Punc_data)
}
```

```{r mapping}
yourMap <- ggplot() +
  geom_polygon(data = tracdata, aes(x=long, y=lat, group = group, fill = Change_Population), colour = "black", size = 0.15)
yourMap
